Bayesian Insights into Aerial Bombing Strategies: An Ordered Logistic Regression Analysis of WWII Target Prioritization Against Germany

### 2024-04-18 paper/paper.qmd
remove unnecessary words and filler phrases

---
title: "Bayesian Insights into Aerial Bombing Strategies: An Ordered Logistic Regression Analysis of WWII Target Prioritization Against Germany"
author: 
  - Yunzhao Li
thanks: "Code and data are available at: https://github.com/yunzhaol/aerial_bomb_priority.git."
date: today
date-format: long
abstract: "During World War II, the strategic decisions about where and how the Allies bombed German industries played a crucial role in the Allied campaign. This study employs Bayesian ordered logistic regression to methodically analyze how different factors—such as the industry type, the launching country, the bomb load, and the number of aircraft—shaped these decisions. Focusing on the preference for targeting crucial oil refineries, the analysis sheds light on the calculated approach to disrupt Germany's war capabilities effectively. The insights from this research not only clarify the strategic rationale behind the Allies' bombing priorities but also enhance our understanding of their impact on the course of the war."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

World War II reshaped global power dynamics, with the strategic aerial bombing campaigns playing a crucial role in determining the outcome of the conflict. The Allied forces’ focused bombardment of German industries was pivotal, aiming to dismantle the economic backbone essential for Germany's war efforts. While considerable research, such as Pape's analysis of strategic bombing's effectiveness [@Pape1996], has been conducted on the impact of these bombings, less attention has been given to the systematic selection process behind targeting specific industrial assets. This paper addresses this understudied aspect by exploring the intricate decision-making process that guided the Allied forces in targeting German industries [@Overy2013].

The crux of this research lies in a detailed examination of a sampled subset of over 5,000 missions from comprehensive wartime records, representing a strategic selection designed to analyze the impact of various factors on target prioritization. These factors include the type of industry targeted, the leading country of the mission, the volume of bombs dropped, and the scale of aircraft involvement. Prior literature, such as studies by Grey [@Grey1999], has often provided broad overviews without delving into the interplay of these specific variables. By applying a Bayesian ordered logistic regression model, this study meticulously quantifies the influence of each factor, offering a clearer picture of the strategic considerations that shaped the Allied bombing strategy.

@@

The paper is organized to provide a clear and detailed examination of the strategic target prioritization during World War II aerial campaigns. Following the introduction in @sec-intro, @sec-data outlines the data employed in the study, describing the sources of historical military data, the variables selected for analysis, and the rationale behind these choices. @sec-model introduces the Bayesian ordered logistic regression model employed to analyze the relationships between our variables of interest, providing a statistical framework for understanding the strategic decisions behind target prioritization during World War II aerial campaigns. @sec-result presents the findings from the Bayesian ordered logistic regression model, detailing how different factors such as target industry and mission country influenced the prioritization of targets. @sec-discussion delves deeper into these results, discussing the strategic implications of the findings, the limitations of the current study, and suggesting avenues for future research to expand upon the insights gained. @sec-model-details shows plots for Posterior predictive check and Diagnostics and their analysis.

@@

The dataset used in this study was sourced from [Data World](https://data.world/datamil/world-war-ii-thor-data), which hosts comprehensive records from the Theater History of Operations (THOR) dataset, documenting aerial bombings during World War II. Originally consisting of approximately 178,282 records, the dataset underwent significant preprocessing to align with the specific focus of this research. The initial step involved filtering out records where the target location was not Germany, reducing the dataset to 64,948 entries. These records form the basis of the analysis, having been retained for detailed examination and modeling of strategic bombing patterns and their impacts.

The data for this study was systematically downloaded, cleaned, analyzed, modeled and visualized using R [@citeR], a comprehensive statistical programming language. The following packages were used for this study。

@@

@fig-industry_country provides a comprehensive view of the processed data, showcasing the distribution of aerial bombardments across different industrial sectors targeted by the Allies during World War II, with a particular focus on those aimed at German industries. The stacked bar chart breaks down the count of missions by the leading country—Great Britain, USA, and others—and prioritizes the targets into four categories: primary target, secondary target, target of opportunity, and target of last resort.

@@

## Comprehensive Understanding of Target Selection {#sec-first-point}

In this study, key variables such as target industry, country flying mission, bomb tons, and aircraft attacking were carefully selected based on their demonstrated strong correlation with target priority. This choice was guided by historical and strategic relevance, ensuring that the model reflects real-world considerations. The quantitative analysis substantiates the critical impact of these variables on strategic military decisions, revealing how different target industries were prioritized based on their role in the war effort. For instance, industries crucial to the enemy's war logistics, such as oil refineries, were often prioritized to cripple the enemy's ability to sustain military operations. This not only aligns with historical records but also provides a granular insight into the decision-making processes, illustrating the careful balancing act between achieving strategic objectives and managing resources effectively during the war.

@@

The scope of this study, while comprehensive, is constrained by the variables selected and the overall dataset size. The omission of variables such as THEATER and TGT_TYPE, although necessary to maintain data quality, limits the depth of strategic analysis possible. Future studies could benefit from incorporating these and other variables, such as TIME_OVER_TARGET and AC_LOST, to provide a fuller picture of the strategic complexities of World War II aerial campaigns.

Moreover, while the model provides a solid foundation, its predictive accuracy could be improved through more sophisticated sampling methods or by integrating additional datasets that offer broader temporal and geographic coverage. For instance, expanding the dataset to include more varied mission types and broader operational theaters could offer new insights into the adaptive strategies employed by the Allies over the course of the war.

Looking ahead, there is significant potential to delve deeper into the effects of bombing strategies on civilian populations by integrating geographical data. This could help in understanding the broader humanitarian impacts of strategic bombing. Additionally, exploring the relationship between flight altitude and mission success could yield valuable insights into the tactical decisions made during these missions. This line of inquiry could benefit from the application of advanced statistical techniques or machine learning models to uncover patterns not readily visible through traditional methods.

@@

Furthermore, understanding the correlation between flying altitude and success rates could shed light on tactical nuances that determined mission outcomes. The wealth of information contained in the dataset paves the way for such multifaceted investigations.

@@

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows a strong influence of the data on the parameter estimates. This plot is crucial for understanding how the evidence in the data alters prior assumptions to shape the final inference drawn from the Bayesian model.

@@

@fig-stanareyouokay-1 is a trace plot. It shows how the sampled values of each parameter evolve over time. Ideally, a well-mixed chain will resemble a 'hairy caterpillar', indicating that the sampling has explored the posterior distribution thoroughly and has likely achieved convergence.

>>>>

---
title: "Bayesian Insights into Aerial Bombing Strategies: An Ordered Logistic Regression Analysis of WWII Target Prioritization Against Germany"
author: 
  - Yunzhao Li
thanks: "Code and data are available at: https://github.com/yunzhaol/aerial_bomb_priority.git."
date: today
date-format: long
abstract: "During World War II, the strategic decisions about where and how the Allies bombed German industries played an essential role in the Allied campaign. This study employs Bayesian ordered logistic regression to methodically analyze how different factors—such as the industry type, the launching country, the bomb load, and the number of aircraft—shaped these decisions. Focusing on the preference for targeting essential oil refineries, the analysis sheds light on the calculated approach to disrupt Germany's war capabilities effectively. The insights from this research not only clarify the strategic rationale behind the Allies' bombing priorities but also enhance our understanding of their impact on the course of the war."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

World War II reshaped global power dynamics, with the strategic aerial bombing campaigns playing a essential role in determining the outcome of the conflict. The Allied forces’ focused bombardment of German industries was pivotal, aiming to dismantle the economic backbone essential for Germany's war efforts. While considerable research, such as Pape's analysis of strategic bombing's effectiveness [@Pape1996], has been conducted on the impact of these bombings, less attention has been given to the systematic selection process behind targeting specific industrial assets. This paper addresses this understudied aspect by exploring the intricate decision-making process that guided the Allied forces in targeting German industries [@Overy2013].

The crux of this research lies in a detailed examination of a sampled subset of over 5,000 missions from extensive wartime records, representing a strategic selection designed to analyze the impact of various factors on target prioritization. These factors include the type of industry targeted, the leading country of the mission, the volume of bombs dropped, and the scale of aircraft involvement. Prior literature, such as studies by Grey [@Grey1999], has often provided broad overviews without delving into the interplay of these specific variables. By applying a Bayesian ordered logistic regression model, this study meticulously quantifies the influence of each factor, offering a clearer picture of the strategic considerations that shaped the Allied bombing strategy.

@@

The paper is organized to provide a clear and detailed examination of the strategic target prioritization during World War II aerial campaigns. Following the introduction in @sec-intro, @sec-data outlines the data employed in the study, describing the sources of historical military data, the variables selected for analysis, and the rationale behind these choices. @sec-model introduces the Bayesian ordered logistic regression model employed to analyze the relationships between our variables of interest, providing a statistical framework for understanding the strategic decisions behind target prioritization during World War II aerial campaigns. @sec-result presents the findings from the Bayesian ordered logistic regression model, detailing how different factors such as target industry and mission country influenced the prioritization of targets. @sec-discussion examines these results, discussing the strategic implications of the findings, the limitations of the current study, and suggesting avenues for future research to expand upon the insights gained. @sec-model-details shows plots for Posterior predictive check and Diagnostics and their analysis.

@@

The dataset used in this study was sourced from [Data World](https://data.world/datamil/world-war-ii-thor-data), which hosts extensive records from the Theater History of Operations (THOR) dataset, documenting aerial bombings during World War II. Originally consisting of approximately 178,282 records, the dataset underwent significant preprocessing to align with the specific focus of this research. The initial step involved filtering out records where the target location was not Germany, reducing the dataset to 64,948 entries. These records form the basis of the analysis, having been retained for detailed examination and modeling of strategic bombing patterns and their impacts.

The data for this study was systematically downloaded, cleaned, analyzed, modeled and visualized using R [@citeR], a extensive statistical programming language. The following packages were used for this study。

@@

@fig-industry_country provides a extensive view of the processed data, showcasing the distribution of aerial bombardments across different industrial sectors targeted by the Allies during World War II, with a particular focus on those aimed at German industries. The stacked bar chart breaks down the count of missions by the leading country—Great Britain, USA, and others—and prioritizes the targets into four categories: primary target, secondary target, target of opportunity, and target of last resort.

@@

## Extensive Understanding of Target Selection {#sec-first-point}

In this study, key variables such as target industry, country flying mission, bomb tons, and aircraft attacking were carefully selected based on their demonstrated strong correlation with target priority. This choice was guided by historical and strategic relevance, ensuring that the model reflects real-world considerations. The quantitative analysis substantiates the critical impact of these variables on strategic military decisions, revealing how different target industries were prioritized based on their role in the war effort. For instance, industries essential to the enemy's war logistics, such as oil refineries, were often prioritized to cripple the enemy's ability to sustain military operations. This not only aligns with historical records but also provides a granular insight into the decision-making processes, illustrating the careful balancing act between achieving strategic objectives and managing resources effectively during the war.

@@

The scope of this study, while extensive, is constrained by the variables selected and the overall dataset size. The omission of variables such as THEATER and TGT_TYPE, although necessary to maintain data quality, limits the depth of strategic analysis possible. Future studies could benefit from incorporating these and other variables, such as TIME_OVER_TARGET and AC_LOST, to provide a fuller picture of the strategic complexities of World War II aerial campaigns.

Moreover, while the model provides a solid foundation, its predictive accuracy could be improved through more advanced sampling methods or by integrating additional datasets that offer broader temporal and geographic coverage. For instance, expanding the dataset to include more varied mission types and broader operational theaters could offer new insights into the adaptive strategies employed by the Allies over the course of the war.

Looking ahead, there is significant potential to examine the effects of bombing strategies on civilian populations by integrating geographical data. This could help in understanding the broader humanitarian impacts of strategic bombing. Additionally, exploring the relationship between flight altitude and mission success could yield valuable insights into the tactical decisions made during these missions. This line of inquiry could benefit from the application of advanced statistical techniques or machine learning models to uncover patterns not readily visible through traditional methods.

@@

Furthermore, understanding the correlation between flying altitude and success rates could clarify tactical nuances that determined mission outcomes. The wealth of information contained in the dataset paves the way for such multifaceted investigations.

@@

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows a strong influence of the data on the parameter estimates. This plot is essential for understanding how the evidence in the data alters prior assumptions to shape the final inference drawn from the Bayesian model.

@@

@fig-stanareyouokay-1 is a trace plot. It shows how the sampled values of each parameter evolve over time. Ideally, a well-mixed chain will resemble a 'hairy caterpillar', indicating that the sampling has explored the posterior distribution extensively and has likely achieved convergence.
